### 2025-04-22 ABP_heatmap_for_Cece_20241018.Rmd
Adding R project, scripts to process data and generate figures, and all helper csv's

---
title: "Cece Sed Traps Manuscript Heat Map"
author: "Adrienne Breef-Pilz"
date: "2024-10-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse, akima, reshape2, gridExtra, colorRamps, RColorBrewer, ggpubr, geomtextpath, scales)

source("https://raw.githubusercontent.com/CareyLabVT/Reservoirs/master/Data/DataAlreadyUploadedToEDI/EDIProductionFiles/Heatmap_EDI_function.R")
```

```{r Read in files}

# Read in from EDI

 EDI_CTD <-
  read_csv("https://pasta.lternet.edu/package/data/eml/edi/200/14/0432a298a90b2b662f26c46071f66b8a")%>%
  #read_csv("./../../Raw_CTD/ctd_L1.csv")|>
   filter(DateTime>ymd_hms("2018-01-01 00:00:00") & DateTime<ymd_hms("2023-12-31 00:00:00"))%>%
   filter(Reservoir=="BVR"|Reservoir=="FCR")%>%
   filter(Site==50)%>%
   select(-starts_with("Flag"))%>%
  mutate(Year=year(DateTime))%>%
  filter(Depth_m>0)

# Make a df of just the casts

CTD_casts <- EDI_CTD |>
  select(Reservoir, DateTime, Year)|>
  unique()

# GEt the month and date for easier graphing. This is not the correct date.   
  CTD_casts$MDay <- format(CTD_casts$DateTime,"%m-%d")
  CTD_casts$MDay <- as.Date(CTD_casts$MDay, "%m-%d")

```

```{r Turnover}

### Still working on this

source("https://raw.githubusercontent.com/LTREB-reservoirs/vera4cast/refs/heads/main/targets/target_functions/target_generation_mixed_binary_daily.R")

bvr_current <- c("https://raw.githubusercontent.com/FLARE-forecast/BVRE-data/bvre-platform-data-qaqc/bvre-waterquality_L1.csv")
#bvr_current = NULL
bvr_historic <- c("https://pasta.lternet.edu/package/data/eml/edi/725/4/9adadd2a7c2319e54227ab31a161ea12")

a <- target_generation_mixed_binary_daily(bvr_current, bvr_historic)

bvr_turnover <- a|>
  mutate(Year = year(datetime),
         DOY = yday(datetime))|>
  filter(observation==1 & DOY>200)|>
  group_by(Year)|>
  slice_min(datetime, n = 1)

# Get turnover for FCR

fcr_historic <- "https://pasta.lternet.edu/package/data/eml/edi/271/8/fbb8c7a0230f4587f1c6e11417fe9dce"

fcr_current <- "https://raw.githubusercontent.com/FLARE-forecast/FCRE-data/refs/heads/fcre-catwalk-data-qaqc/fcre-waterquality_L1.csv"

ag <- target_generation_mixed_binary_daily(fcr_current, fcr_historic)

fcr_turnover <- ag|>
  mutate(Year = year(datetime),
         DOY = yday(datetime))|>
  filter(observation==1 & DOY>200)|>
  group_by(Year)|>
  slice_min(datetime, n = 1)

## in the dates Cece has here

turnover <- dplyr::bind_rows(fcr_turnover,bvr_turnover)|>
  mutate(Reservoir=ifelse(site_id=="fcre", "FCR", "BVR"))|>
  filter(datetime<as.Date("2024-01-01"))|>
  select(datetime, Reservoir, Year)

# Add in BVR observations
turn <- data.frame(
  datetime = c(as.Date("2018-10-24"), as.Date("2019-11-02")),
  Reservoir = c("BVR", "BVR"),
  Year = c(2018, 2019)
)

turnover <- dplyr::bind_rows(turn, turnover)

# GEt the month and date for easier graphing. This is not the correct date.   
  turnover$MDay <- format(turnover$datetime,"%m-%d")
  turnover$MDay <- as.Date(turnover$MDay, "%m-%d")
```

```{r make heat map df}

# This function makes a data frame of interpolated observations over the years for making multiple heat maps using facet_grid

heat_interp <- function(
    data,
    var,
    Res,
    Years){

  # data= "EDI_CTD_temp"
  # var = "Temp_C"
  # Res = "FCR"
  # Years = 2023

  interp <- data.frame(x = NA, y = NA, z = NA, Reservoir = NA)

 bvr <- get(data) %>%
   select(Reservoir, DateTime, Depth_m, var) |>
        mutate(Year = year(DateTime),
               Date = as.Date(DateTime))%>%
        filter(#Reservoir == "BVR",
               !is.na(get(var)),
               !is.na(Depth_m),
               Reservoir %in% Res,
               Year %in% Years)

      if(nrow(bvr)>0){
        interp <- interp2xyz(interp(x = as.numeric(bvr$Date), y= bvr$Depth_m, z= unlist(bvr[[var]]),
                                        xo = seq(min(bvr$Date), max(bvr$Date), "day"),
                                        yo = seq(min(bvr$Depth_m), max(bvr$Depth_m), 
                                                 by = 0.1), 
                                        duplicate = "mean"),data.frame = T)
        interp2<-interp%>%
    mutate(Date=as.Date(x, origin = "1970-01-01"),
           Reservoir=Res)

  return(interp2)
  }
}
```

```{r plotting}

#custom_heatmap_palettes <- c("#313695", "#4575b4", "#74add1",
#                            "#abd9e9", "#e0f3f8", "#ffffbf", 
#                            "#fee090","#fdae61", "#f46d43", 
#                            "#d73027", "#a50026")

make_heat_map <- function(
    data,
    sample_dates,
    turnovers, 
    y_lim,
    custom_heatmap_palette,
    y_units,
    output_file){

#Heatmap

# change the plots dimensions  
jpeg(paste0(output_file),width = 8, height = 4, units = "in", res = 400)
water_temp <- data%>%
  mutate(x = as.Date(x, origin = "1970-01-01"),
         Year = year(x))%>%
  filter(!is.na(z))%>%
  ggplot()+
  geom_raster(aes(x=MDay, y=y,fill = z))+
  geom_point(data = sample_dates, aes(x = MDay, y = 0.1, z = NULL), pch = '|', size = 2, color = "black") + #to mark cast dates 
   geom_vline(data= turnovers, aes(xintercept=MDay) , linetype="solid", color="black", linewidth=0.5) +
  labs(y = "Depth (m)") +
  scale_x_date(expand = c(0,0)) +
  scale_y_reverse(expand = c(0,0)) +
  #scale_y_continuous(expand = c(0,0))+
  theme(panel.border = element_rect(fill = NA),
        axis.title.x=element_blank())+
  scale_fill_gradientn(name = y_units,
                       limits = y_lim,
                       colours = custom_heatmap_palette)+
  facet_grid(rows= vars(Reservoir),cols = vars(Year),
             scales = "free_y",
             space = "free"
             )+
  theme(panel.background = element_rect(fill = "white"),
          panel.spacing.x = unit(0.05, "lines"),
        legend.position = 'bottom',
        panel.grid = element_blank())
  theme_bw(base_size = 18)
  # labs(color = "")+
  # geom_point(data = sensor_depth_plot, 
  #            aes(x = Date, y = Sensor_depth), 
  #            shape = 23, fill = "black", color = "white", size  =2.5)+
  # geom_vline(data=lines,aes(xintercept = Dates), show.legend = F)+
  # scale_linetype_manual(values = c("solid","84"),
  #                       limits = c("Thermocline","Outtake pipe"),  
  #                       name= NULL)+

print(water_temp)
dev.off()

}
```

```{r Make Temp plot}

# EDI_CTD_temp <- EDI_CTD|>
#   select(Reservoir, DateTime, Year, Depth_m, Temp_C)

wer <- data.frame(
  data= c(rep("EDI_CTD",12)),
  var = c(rep("Temp_C",12)),
  Res = c(rep("FCR",6), rep("BVR",6)),
  Years = c(rep(2018:2023,2))
)

system.time(sd <- wer%>%
             pmap(heat_interp))

#Bind all the QAQC files from above
   temp_df <- dplyr::bind_rows(sd)

  temp_df <- temp_df%>%
    filter(!is.na(Reservoir),
           y>0.1)

# GEt the month and date for easier graphing. This is not the correct date.   
  temp_df$MDay <- format(temp_df$Date,"%m-%d")
temp_df$MDay <- as.Date(temp_df$MDay, "%m-%d")

# make the plot

make_heat_map(
  data= temp_df,
  sample_dates = CTD_casts,
  turnovers = turnover, 
  y_lim = c(-1, 30),
  custom_heatmap_palette = blue2green2red(60),
  y_units = "Water temperature (°C)",
  output_file = "figures/SedTraps_CTD_waterTemp_2018_2023.jpeg"
)

```

```{r Make DO plot}

# EDI_CTD_temp <- EDI_CTD|>
#   select(Reservoir, DateTime, Year, Depth_m, Temp_C)

wer <- data.frame(
  data= c(rep("EDI_CTD",12)),
  var = c(rep("DO_mgL",12)),
  Res = c(rep("FCR",6), rep("BVR",6)),
  Years = c(rep(2018:2023,2))
)

system.time(sd <- wer%>%
             pmap(heat_interp))

#Bind all the QAQC files from above
   temp_df <- dplyr::bind_rows(sd)

  temp_df <- temp_df%>%
    filter(!is.na(Reservoir),
           y>0.1)

# GEt the month and date for easier graphing. This is not the correct date.   
  temp_df$MDay <- format(temp_df$Date,"%m-%d")
temp_df$MDay <- as.Date(temp_df$MDay, "%m-%d")

# make the plot

make_heat_map(
  data= temp_df,
  sample_dates = CTD_casts,
  turnovers = turnover, 
  y_lim = c(-1,16),
  custom_heatmap_palette = rev(blue2green2red(60)),
  y_units = "Dissolved Oxygen mg/L",
  output_file = "figures/SedTraps_CTD_DO_2018_2023.jpeg"
)

```

```{r Need to add in handhelp pH obs}

ph <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/198/12/e0181372c6d4cbc66eca4644f8385470" )|>
  #read_csv("./../../../DataAlreadyUploadedToEDI/EDIProductionFiles/MakeEMLYSI_PAR_secchi/2023/YSI_PAR_profiles_2013_2023.csv")|>
  filter(DateTime>as.Date("2018-01-01"))|>
  filter(Reservoir=="BVR"|Reservoir=="FCR")|>
   filter(Site==50)|>
  select(Reservoir, DateTime, Depth_m, pH)|>
  drop_na(pH)|>
  mutate(Year=year(DateTime))

# bind with the CTD

EDI_CTD_ph <- dplyr::bind_rows(EDI_CTD,ph)

CTD_casts_ph <- ph|>
  select(Reservoir, DateTime, Year)|>
  mutate(DateTime = as.Date(DateTime))|>
  unique()%>%
  dplyr::bind_rows(. , CTD_casts)

# GEt the month and date for easier graphing. This is not the correct date.   
  CTD_casts_ph$MDay <- format(CTD_casts_ph$DateTime,"%m-%d")
  CTD_casts_ph$MDay <- as.Date(CTD_casts_ph$MDay, "%m-%d")
```

```{r Make pH plot}

##Need to fix this

 EDI_CTD_pH <- EDI_CTD|>
   select(Reservoir, DateTime, Year, Depth_m, pH)%>%
  bind_rows(.,ph)

wer <- data.frame(
  data= c(rep("EDI_CTD_pH",12)),
  var = c(rep("pH",12)),
  Res = c(rep("FCR",6), rep("BVR",6)),
  Years = c(rep(2018:2023,2))
)

system.time(sd <- wer%>%
             pmap(heat_interp))

#Bind all the QAQC files from above
   temp_df <- dplyr::bind_rows(sd)

  temp_df <- temp_df%>%
    filter(!is.na(Reservoir),
           y>0.1)

# GEt the month and date for easier graphing. This is not the correct date.   
  temp_df$MDay <- format(temp_df$Date,"%m-%d")
temp_df$MDay <- as.Date(temp_df$MDay, "%m-%d")

# make the plot

make_heat_map(
  data= temp_df,
  sample_dates = CTD_casts_ph,
  turnovers = turnover,
  y_lim = c(5,10),
  custom_heatmap_palette = blue2green2red(60),
  y_units = "pH",
  output_file = "figures/SedTraps_CTD_pH_2018_2023.jpeg"
)

```

```{r metals plotting}

make_metals_heat_map <- function(
    data,
    sample_dates,
    turnovers, 
    y_lim,
    custom_heatmap_palette,
    y_units){

#Heatmap

# change the plots dimensions  
#jpeg(paste0(output_file),width = 8, height = 4, units = "in", res = 400)
water_temp <- data%>%
  mutate(x = as.Date(x, origin = "1970-01-01"),
         Year = year(x))%>%
  filter(!is.na(z))%>%
  ggplot()+
  geom_raster(aes(x=MDay, y=y,fill = z))+
  geom_point(data = sample_dates, aes(x = MDay, y = 0.1, z = NULL), pch = '|', size = 2, color = "gray80") + #to mark cast dates 
   geom_vline(data= turnovers, aes(xintercept=MDay) , linetype="solid", color="gray80", linewidth=0.5) +
  labs(y = "Depth (m)") +
  scale_x_date(expand = c(0,0)) +
  scale_y_reverse(expand = c(0,0),
                  breaks = c(0.0, 2.5, 5.0, 7.5)) +
  #scale_y_continuous(expand = c(0,0))+
  theme(panel.border = element_rect(fill = NA),
        axis.title.x=element_blank()) +
  scale_fill_gradientn(name = y_units,
                       limits = y_lim,
                       colours = custom_heatmap_palette,
                       values=rescale(c(0,5,20)))+
  facet_grid(rows= vars(Reservoir),cols = vars(Year),
             scales = "free_y",
             space = "free"
             )+
  theme(panel.background = element_rect(fill = "white"),
          panel.spacing = unit(0.05, "lines"),
        legend.position = 'bottom',
        panel.grid = element_blank())
  theme_bw(base_size = 18)
  # labs(color = "")+
  # geom_point(data = sensor_depth_plot, 
  #            aes(x = Date, y = Sensor_depth), 
  #            shape = 23, fill = "black", color = "white", size  =2.5)+
  # geom_vline(data=lines,aes(xintercept = Dates), show.legend = F)+
  # scale_linetype_manual(values = c("solid","84"),
  #                       limits = c("Thermocline","Outtake pipe"),  
  #                       name= NULL)+

return(water_temp)
dev.off()

}
```

```{r}
metals <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/455/8/9c8c61b003923f4f03ebfe55cea8bbfd")

# filter the data to the time you want. Right now this is just for 2021.

met <- metals |>
  filter(year(DateTime)>= 2018 & Depth_m <= 9)

met <- met %>% 
  filter(Flag_TFe_mgL != 8 & Flag_SFe_mgL != 8 & Flag_TFe_mgL != 68 & Flag_SFe_mgL != 68,
       Flag_TMn_mgL != 8 & Flag_SMn_mgL != 8 & Flag_TMn_mgL != 68 & Flag_SMn_mgL != 68)

# ratios of S to T for plotting 
met <- met %>% 
  mutate(FeDiff = TFe_mgL - SFe_mgL,
         MnDiff = TMn_mgL - SMn_mgL,
         Year = year(DateTime),
         Date = format(DateTime,"%m-%d"))

#create a data frame with the correct values
metratio <- met %>% 
  filter(FeDiff > 0,
         MnDiff > 0)

#create a data frame with the incorrect values and switch them
met_issues <- met %>% 
  filter(FeDiff < 0,
         MnDiff < 0) %>% 
  rename('test1' = 'TMn_mgL',
         'test2' = 'SMn_mgL',
         'testA' = 'TFe_mgL',
         'testB' = 'SFe_mgL') %>% 
  rename('TMn_mgL' = 'test2',
         'SMn_mgL' = 'test1',
         'TFe_mgL' = 'testB',
         'SFe_mgL' = 'testA') %>% 
  mutate(MnDiff = TMn_mgL - SMn_mgL,
         FeDiff = TFe_mgL - SFe_mgL)

#rbind to rejoin
metratio <- rbind(metratio, met_issues) %>% 
  mutate(Fe_ratio = SFe_mgL/TFe_mgL,
         Mn_ratio = SMn_mgL/TMn_mgL)

metDates <- met %>% 
  select(Reservoir, DateTime, Year) %>% 
  unique()

metDates$MDay <- format(metDates$DateTime,"%m-%d")
  metDates$MDay <- as.Date(metDates$MDay, "%m-%d")
```

```{r Make TMn plot}
wer <- data.frame(
  data= c(rep("met",12)),
  var = c(rep("TMn_mgL",12)),
  Res = c(rep("FCR",6), rep("BVR",6)),
  Years = c(rep(2018:2023,2))
)

system.time(sd <- wer%>%
             pmap(heat_interp))

#Bind all the QAQC files from above
   temp_df <- dplyr::bind_rows(sd)

  temp_df <- temp_df%>%
    filter(!is.na(Reservoir),
           y>0.1)

# GEt the month and date for easier graphing. This is not the correct date.   
temp_df$MDay <- format(temp_df$Date,"%m-%d")
temp_df$MDay <- as.Date(temp_df$MDay, "%m-%d")

# make the plot

p1 <- make_metals_heat_map(
  data= temp_df,
  sample_dates = metDates,
  turnovers = turnover, 
  y_lim = c(0,2.5),
  custom_heatmap_palette = blue2green2red(60),
  y_units = "mg/L"
) + ggtitle('Total Mn')

```

```{r Make SMn plot}
wer <- data.frame(
  data= c(rep("met",12)),
  var = c(rep("SMn_mgL",12)),
  Res = c(rep("FCR",6), rep("BVR",6)),
  Years = c(rep(2018:2023,2))
)

system.time(sd <- wer%>%
             pmap(heat_interp))

#Bind all the QAQC files from above
   temp_df <- dplyr::bind_rows(sd)

  temp_df <- temp_df%>%
    filter(!is.na(Reservoir),
           y>0.1)

# GEt the month and date for easier graphing. This is not the correct date.   
temp_df$MDay <- format(temp_df$Date,"%m-%d")
temp_df$MDay <- as.Date(temp_df$MDay, "%m-%d")

# make the plot

p2 <- make_metals_heat_map(
  data= temp_df,
  sample_dates = metDates,
  turnovers = turnover, 
  y_lim = c(0,2.5),
  custom_heatmap_palette = blue2green2red(60),
  y_units = "mg/L"
) + ggtitle('Soluble Mn')

```

```{r}
MnPlot <- ggarrange(p1, p2, ncol = 1, labels = c('A', 'B'), common.legend = TRUE,
                    legend = 'bottom')
MnPlot
ggsave(MnPlot, filename = "figures/SedTraps_Mn_2018_2023.png", height = 6, width = 8, units = 'in')
```

```{r Make TFe plot}
wer <- data.frame(
  data= c(rep("met",12)),
  var = c(rep("TFe_mgL",12)),
  Res = c(rep("FCR",6), rep("BVR",6)),
  Years = c(rep(2018:2023,2))
)

system.time(sd <- wer%>%
             pmap(heat_interp))

#Bind all the QAQC files from above
   temp_df <- dplyr::bind_rows(sd)

  temp_df <- temp_df%>%
    filter(!is.na(Reservoir),
           y>0.1)

# GEt the month and date for easier graphing. This is not the correct date.   
temp_df$MDay <- format(temp_df$Date,"%m-%d")
temp_df$MDay <- as.Date(temp_df$MDay, "%m-%d")

# make the plot

p3 <- make_metals_heat_map(
  data= temp_df,
  sample_dates = metDates,
  turnovers = turnover, 
  y_lim = c(0,20),
  custom_heatmap_palette = blue2green2red(60),
  y_units = "mg/L"
) + ggtitle('Total Fe')

```

```{r Make SFe plot}
wer <- data.frame(
  data= c(rep("met",12)),
  var = c(rep("SFe_mgL",12)),
  Res = c(rep("FCR",6), rep("BVR",6)),
  Years = c(rep(2018:2023,2))
)

system.time(sd <- wer%>%
             pmap(heat_interp))

#Bind all the QAQC files from above
   temp_df <- dplyr::bind_rows(sd)

  temp_df <- temp_df%>%
    filter(!is.na(Reservoir),
           y>0.1)

# GEt the month and date for easier graphing. This is not the correct date.   
temp_df$MDay <- format(temp_df$Date,"%m-%d")
temp_df$MDay <- as.Date(temp_df$MDay, "%m-%d")

# make the plot

p4 <- make_metals_heat_map(
  data= temp_df,
  sample_dates = metDates,
  turnovers = turnover, 
  y_lim = c(0,20),
  custom_heatmap_palette = blue2green2red(60),
  y_units = "mg/L"
) + ggtitle('Soluble Fe')

```

```{r}
FePlot <- ggarrange(p3, p4, ncol = 1, labels = c('A', 'B'), common.legend = TRUE,
                    legend = 'bottom')
FePlot
ggsave(FePlot, filename = "figures/SedTraps_Fe_2018_2023.png", height = 6, width = 8, units = 'in')
```