### 2025-05-06 Manuscript1_figures.Rmd

---
title: "Manuscript1_figures"
author: "CEW"
date: "2024-11-14"
output: 
  html_document:
  df_print: paged

---

## Data sources:

1. Schreiber, M.E., C.E. Wood, A. Breef-Pilz, C.E. Bauer, G. Haynie, R. Keverline, M. Kricheldorf, and E. Tipper. 2024. Time series of total and soluble metal concentrations from Falling Creek Reservoir, Beaverdam Reservoir, and Carvins Cove Reservoir in southwestern Virginia, USA from 2014 through 2023 ver 8. Environmental Data Initiative. https://doi.org/10.6073/pasta/584b1daf7f1541bc595aa3a9dd71a2c7.

2. Schreiber, M.E., C.E. Wood, C.E. Bauer, A. Breef-Pilz, G.K. Haynie, and E.M. Tipper. 2024. Sediment trap time series data for Beaverdam Reservoir and Falling Creek Reservoir in southwestern Virginia, USA 2018 through 2023 ver 8. Environmental Data Initiative. https://doi.org/10.6073/pasta/5ee061465e62c29587042c58e7a91e13.

3. Carey, C.C., A. Breef-Pilz, V. Daneshmand, A.D. Delany, and R.Q. Thomas. 2024. Time series of high-frequency sensor data measuring water temperature, dissolved oxygen, pressure, conductivity, specific conductance, total dissolved solids, chlorophyll a, phycocyanin, fluorescent dissolved organic matter, and turbidity at discrete depths in Falling Creek Reservoir, Virginia, USA in 2018-2023 ver 8. Environmental Data Initiative. https://doi.org/10.6073/pasta/7541e8d297850be7c613d116156735a9.

4. Carey, C.C. and A. Breef-Pilz. 2025. Time series of high-frequency sensor data measuring water temperature, dissolved oxygen, conductivity, specific conductance, total dissolved solids, chlorophyll a, phycocyanin, fluorescent dissolved organic matter, and turbidity at discrete depths, and water level in Beaverdam Reservoir, Virginia, USA in 2009-2024 ver 5. Environmental Data Initiative. https://doi.org/10.6073/pasta/8f666b34c120aa5d2242964cf3147f90.

5. Carey, C.C., A. Breef-Pilz, H.L. Wander, D.W. Howard, G.K. Haynie, M.K. Kricheldorf, S. Tannheiser, and E.M. Tipper. 2025. Secchi depth data and discrete depth profiles of water temperature, dissolved oxygen, conductivity, specific conductance, photosynthetic active radiation, oxidation-reduction potential, and pH for Beaverdam Reservoir, Carvins Cove Reservoir, Falling Creek Reservoir, Gatewood Reservoir, and Spring Hollow Reservoir in southwestern Virginia, USA 2013-2024 ver 13. Environmental Data Initiative. https://doi.org/10.6073/pasta/0df2b12c947e52a8d0240a75d20bbc1e.

## Figures 6-9, S1, S3, and S4
```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(pacman)
pacman::p_load(tidyverse, readxl, stringr, readr, glmnet, ISLR, lubridate, ggplot2, ggpubr, stats, gridExtra, graphics, gt)
```

```{r pull in all data, eval=TRUE, include=FALSE}
metals_file <- read_csv('https://pasta.lternet.edu/package/data/eml/edi/455/8/9c8c61b003923f4f03ebfe55cea8bbfd')

fluxes <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/1474/8/7cc135d14d92911eb7e8f6ceb096a29d")

FCRDO <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/271/8/fbb8c7a0230f4587f1c6e11417fe9dce")

BVRDO1 <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/725/5/f649de0e8a468922b40dcfa34285055e")

BVRDO2 <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/725/5/b7f0524aee78dcf4a2452a203a39c1f2")

Handhelds  <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/198/13/e50a50d062ee73f4d85e4f20b360ce4f")
```

```{r set up file path}
#set up file path for figures, change if necessary
myFilePath <- './figures'
```

## Figure 6. DO at HOx Activation vs Max Fe and Mn concentrations

```{r DO plot figure 6 script, eval=TRUE, include=FALSE}

#First HOx activation dates, where HOx was activated for more than 7 days before deactivation, see Table 1 in manuscript

HOxBoxes <- read_csv('HOx_Ranges.csv') %>% 
  mutate(Group = rownames(.),
         Layer = 'Hypolimnion',
         Year = year(HOxOn),
         DOYon = yday(HOxOn),
         DOYoff = yday(HOxOff))

HOxOn <- HOxBoxes %>%
  mutate(Duration = HOxOff - HOxOn) %>% 
  filter(Duration >= 7) %>% 
  select(Reservoir, Year, HOxOn) %>% 
  group_by(Reservoir, Year) %>% 
  slice_min(order_by = HOxOn)

DOps1 <- FCRDO %>%
  select(Reservoir, Site, DateTime, RDOsat_percent_9_adjusted) %>% 
  mutate(Date = date(DateTime)) %>%
  group_by(Reservoir, Site, Date) %>% 
  summarise(AvgDOsat_percent_9 = mean(RDOsat_percent_9_adjusted, na.rm = TRUE)) %>% 
  ungroup() %>% #now select first HOx activation dates
  filter(Date == as.Date('2018-04-23') | Date == as.Date('2019-06-03') | 
           Date == as.Date('2020-06-29') | Date == as.Date('2021-06-11') |
           Date == as.Date('2022-05-19') | Date == as.Date('2023-03-06')) %>% 
  mutate(Year = year(Date)) %>% 
  rename('HOxStart' = 'Date')

DO_Handhelds <- Handhelds %>% 
  filter(Reservoir == 'FCR') %>% 
  mutate(Date = date(DateTime),
         Year = year(DateTime)) %>% #HOx activated on 2018-04-23 @ noon
  filter(DateTime == as_datetime('2018-04-23 09:45:00') & Depth_m == 9) %>% 
  select(Reservoir, Site, Date, Year, DOsat_percent) %>% 
  rename('HOxStart' = 'Date',
         'AvgDOsat_percent_9' = 'DOsat_percent')

DOps2 <- DOps1 %>% 
  full_join(DO_Handhelds)

metals <- metals_file %>% 
  filter(metals_file$DateTime >= as.Date('2018-01-01'),
         Site == 50, Reservoir == 'FCR' | Reservoir == 'BVR') %>% 
  select(Reservoir, Site, DateTime, Depth_m, TFe_mgL, SFe_mgL, TMn_mgL, SMn_mgL,
         Flag_TFe_mgL, Flag_TMn_mgL, Flag_SFe_mgL, Flag_SMn_mgL) %>%
  filter(Flag_TFe_mgL != 8 & Flag_SFe_mgL != 8 & Flag_TFe_mgL != 68 & Flag_SFe_mgL != 68,
         Flag_TMn_mgL != 8 & Flag_SMn_mgL != 8 & Flag_TMn_mgL != 68 & Flag_SMn_mgL != 68) %>% 
  mutate(FeDiff = TFe_mgL - SFe_mgL,
         MnDiff = TMn_mgL - SMn_mgL,
         STFe = SFe_mgL/TFe_mgL,
         STMn = SMn_mgL/TMn_mgL,
         Date = date(DateTime),
         Layer = NA,
         Layer = ifelse(Reservoir == 'FCR' & Depth_m <= 4, 'Epilimnion', Layer),
         Layer = ifelse(Reservoir == 'FCR' & Depth_m >= 4, 'Hypolimnion', Layer),
         Layer = ifelse(Reservoir == 'BVR' & Depth_m <= 5, 'Epilimnion', Layer),
         Layer = ifelse(Reservoir == 'BVR' & Depth_m >= 5, 'Hypolimnion', Layer),
         Year = year(DateTime)) %>% 
  group_by(Reservoir, Site, Date, Layer) %>% 
  mutate(AvgSTFe = mean(STFe, na.rm = TRUE),
         AvgSTMn = mean(STMn, na.rm = TRUE)) %>% 
  filter(FeDiff >= -0.02,
         MnDiff >= -0.06)

MaxMetals <- metals %>% 
  filter(Year >= 2018 & Depth_m == 9 & Reservoir == 'FCR') %>% 
  filter(Year == 2018 & between(Date, as.Date('2018-04-23'), as.Date('2018-08-12')) | #only include dates where HOx
           Year == 2019 & between(Date, as.Date('2019-06-03'), as.Date('2019-11-20'))  | #was active
           Year == 2020 & between(Date, as.Date('2020-06-29'), as.Date('2020-12-02')) |
           Year == 2021 & between(Date, as.Date('2021-06-11'), as.Date('2021-12-06')) |
           Year == 2022 & between(Date, as.Date('2022-05-19'), as.Date('2022-12-12')) |
           Year == 2023 & between(Date, as.Date('2023-03-06'), as.Date('2023-12-04'))) %>% 
  group_by(Reservoir, Site, Year) %>% 
  summarise(MaxTFe_mgL = max(TFe_mgL),
            MaxTMn_mgL = max(TMn_mgL))

DO_MaxMetalMass <- DOps2 %>% 
  full_join(MaxMetals)

DO_MaxMetal_Feplot <- DO_MaxMetalMass %>% 
  ggplot() + theme_bw(base_size = 18) +
  geom_point(aes(x = AvgDOsat_percent_9, y = MaxTFe_mgL, color = as.factor(Year),
                 shape = as.factor(Year)), size = 3) +
  ylab('Maximum Concentration (mg/L)') + xlab('') +
  guides(color = guide_legend(title = 'Year'), shape = guide_legend(title = 'Year')) +
  ggtitle('Total Iron')

DO_MaxMetal_Mnplot <- DO_MaxMetalMass %>% 
  ggplot() + theme_bw(base_size = 18) +
  geom_point(aes(x = AvgDOsat_percent_9, y = MaxTMn_mgL, color = as.factor(Year),
                 shape = as.factor(Year)), size = 3) +
  ylab('') + xlab('') +
  guides(color = guide_legend(title = 'Year'), shape = guide_legend(title = 'Year')) +
  ggtitle('Total Manganese')

DO_MaxMetal_plot <- ggarrange(DO_MaxMetal_Feplot, DO_MaxMetal_Mnplot, nrow = 1, common.legend = TRUE,
                              legend = 'right', labels = c('A', 'B'))
DO_MaxMetal_plotFinal <- annotate_figure(DO_MaxMetal_plot, bottom = text_grob('Dissolved Oxygen at HOx Activation (% saturation)', size = 18, hjust = 0.53, vjust = -1))
```

```{r Figure 6, eval=TRUE, include=TRUE}
DO_MaxMetal_plotFinal
```

## Figure 7. Seasonal Flux Paired T-Test Plot

```{r Fig 7 and 8 set up, eval=TRUE, include=FALSE}
 fluxes2018_2023 <-  fluxes %>% 
  select(Reservoir, Date, Depth_m, Duration_days, TFeFlux_gm2d, TMnFlux_gm2d) %>% 
  na.omit() %>%
  mutate(Layer = if_else(Depth_m <= 5, 'Epilimnion', 'Hypolimnion'),
         Year = year(Date))

avgFluxes2018_2023 <- fluxes2018_2023 %>% 
  group_by(Reservoir, Date, Depth_m, Layer, Year) %>% 
  summarise(Duration_days = mean(Duration_days),
            AvgTFeFlux_gm2d = mean(TFeFlux_gm2d, na.rm = TRUE),
            AvgTMnFlux_gm2d = mean(TMnFlux_gm2d, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(TO = NA,
         #2018, only include data points from FCR that match data points from BVR
         TO = if_else(Reservoir == 'FCR' & Layer == 'Epilimnion' 
                      & Date < as.Date('2018-09-03'), 'Pre_', TO),
         TO = if_else(Reservoir == 'FCR' & Layer == 'Hypolimnion' 
                      & Date < as.Date('2018-07-23'), 'Pre_', TO),
         TO = if_else(Reservoir == 'FCR' & Layer == 'Epilimnion' 
                      & between(Date, as.Date('2018-09-03'), as.Date('2018-10-21')), 'Pre', TO),
         TO = if_else(Reservoir == 'FCR' & Layer == 'Hypolimnion' 
                      & between(Date, as.Date('2018-07-23'), as.Date('2018-10-21')), 'Pre', TO),
         TO = if_else(Year == 2018 & Reservoir == 'BVR' 
                      & Date <= as.Date('2018-09-07'), 'Pre', TO),
         TO = if_else(Year == 2018 & Reservoir == 'FCR' 
                      & Date >= as.Date('2018-10-21'), 'Post', TO),
         TO = if_else(Year == 2018 & Reservoir == 'BVR' 
                      & Date >= as.Date('2018-10-21'), 'Post', TO),
         #2019
         TO = if_else(Year == 2019 & Reservoir == 'FCR' & Date < as.Date('2019-10-24'), 'Pre', TO),
         TO = if_else(Year == 2019 & Reservoir == 'BVR' & Date < as.Date('2019-10-30'), 'Pre', TO),
         TO = if_else(Year == 2019 & Reservoir == 'FCR' & Date >= as.Date('2019-10-24'), 'Post', TO),
         TO = if_else(Year == 2019 & Reservoir == 'BVR' & Date >= as.Date('2019-10-30'), 'Post', TO),
         #2020
         TO = ifelse(Year == 2020 & Reservoir == 'FCR' & Date < as.Date('2020-11-02'), 'Pre', TO),
         TO = ifelse(Year == 2020 & Reservoir == 'BVR' & Date < as.Date('2020-11-02'), 'Pre', TO),
         TO = ifelse(Year == 2020 & Reservoir == 'FCR' & Date >= as.Date('2020-11-02'), 'Post', TO),
         TO = ifelse(Year == 2020 & Reservoir == 'BVR' & Date >= as.Date('2020-11-02'), 'Post', TO),
         #2021
         TO = ifelse(Year == 2021 & Reservoir == 'FCR' & Date < as.Date('2021-11-03'), 'Pre', TO),
         TO = ifelse(Year == 2021 & Reservoir == 'BVR' & Date < as.Date('2021-11-05'), 'Pre', TO),
         TO = ifelse(Year == 2021 & Reservoir == 'FCR' & Date >= as.Date('2021-11-03'), 'Post', TO),
         TO = ifelse(Year == 2021 & Reservoir == 'BVR' & Date >= as.Date('2021-11-05'), 'Post', TO),
         #2022
         TO = ifelse(Year == 2022 & Reservoir == 'FCR' & Date < as.Date('2022-10-20'), 'Pre', TO),
         TO = ifelse(Year == 2022 & Reservoir == 'BVR' & Date < as.Date('2022-10-19'), 'Pre', TO),
         TO = ifelse(Year == 2022 & Reservoir == 'FCR' & Date >= as.Date('2022-10-20'), 'Post', TO),
         TO = ifelse(Year == 2022 & Reservoir == 'BVR' & Date >= as.Date('2022-10-19'), 'Post', TO),
         #2023
         TO = ifelse(Year == 2023 & Reservoir == 'FCR' & Date < as.Date('2023-11-02'), 'Pre', TO),
         TO = ifelse(Year == 2023 & Reservoir == 'BVR' & Date < as.Date('2023-11-02'), 'Pre', TO),
         TO = ifelse(Year == 2023 & Reservoir == 'FCR' & Date >= as.Date('2023-11-02'), 'Post', TO),
         TO = ifelse(Year == 2023 & Reservoir == 'BVR' & Date >= as.Date('2023-11-02'), 'Post', TO)) %>% 
  arrange(Date)

avgCumulFluxes2018_2023 <- avgFluxes2018_2023 %>% 
  filter(TO == 'Pre' & Layer == 'Hypolimnion') %>%
  group_by(Reservoir, Layer, Year) %>% 
  mutate(CumulFe_gm2 = cumsum(AvgTFeFlux_gm2d*Duration_days),
         CumulMn_gm2 = cumsum(AvgTMnFlux_gm2d*Duration_days),
         CumulDuration_days = cumsum(Duration_days))

CumulFluxSummary <- avgCumulFluxes2018_2023 %>%
  group_by(Reservoir, Layer, Year) %>% 
  summarise(CumulFe_gm2 = max(CumulFe_gm2),
            CumulMn_gm2 = max(CumulMn_gm2),
            Duration_days = sum(Duration_days),
            Date = max(Date)) %>% 
  mutate(FeFlux = round(CumulFe_gm2/Duration_days, digits = 3),
         MnFlux = round(CumulMn_gm2/Duration_days, digits = 3))

avgCumulFluxes2018_2023_long <- avgCumulFluxes2018_2023 %>% 
  select(-c(AvgTFeFlux_gm2d, AvgTMnFlux_gm2d)) %>% 
  pivot_longer(cols = 8:9, values_to = 'Fluxes_gm2', names_to = 'Metal')

CumulFluxSummary_long <- CumulFluxSummary %>% 
  pivot_longer(cols = 4:5, values_to = 'Fluxes_gm2', names_to = 'Metal') %>% 
  mutate(Metal = factor(Metal, levels = c('CumulFe_gm2', 'CumulMn_gm2'),
                        labels = c('Iron', 'Manganese')))
```

```{r Figure 7 script, eval=TRUE, include=FALSE}
CumulFePlot1 <- CumulFluxSummary %>% 
  ggplot() + theme_bw(base_size = 18) + ggtitle('Iron') + ylab('Fe Flux (g/m2d)') +
  geom_point(aes(x = Reservoir, y = FeFlux, color = as.factor(Year), shape = as.factor(Year)), size = 3) +
  #2018
  geom_segment(x = 'BVR', xend = 'FCR', y = 0.209, yend = 1.196, alpha = 0.5,
               color = "#F8766D", linetype = 3) +
  #2019
  geom_segment(x = 'BVR', xend = 'FCR', y = 0.661, yend = 1.398, alpha = 0.5,
               color = "#B79F00", linetype = 3) +
  #2020
  geom_segment(x = 'BVR', xend = 'FCR', y = 0.276, yend = 1.676, alpha = 0.5,
               color = "#00BA38", linetype = 3) +
  #2021
  geom_segment(x = 'BVR', xend = 'FCR', y = 0.775, yend = 1.776, alpha = 0.5,
               color = "#00BFC4", linetype = 3) +
  #2022
  geom_segment(x = 'BVR', xend = 'FCR', y = 0.867, yend = 1.849, alpha = 0.5,
               color = "#619CFF", linetype = 3) +
  #2023
  geom_segment(x = 'BVR', xend = 'FCR', y = 0.819, yend = 1.017, alpha = 0.5,
               color = "#F564E3", linetype = 3) +
  annotate(geom = 'text', x= 1.5, y = 1.8, label = '**', size = 12) +
  guides(color = guide_legend(title = 'Year'), shape = guide_legend(title = 'Year'))

CumulMnPlot1 <- CumulFluxSummary %>% 
  ggplot() + theme_bw(base_size = 18) + ggtitle('Manganese') + ylab('Mn Flux (g/m2d)') +
  geom_point(aes(x = Reservoir, y = MnFlux, color = as.factor(Year), shape = as.factor(Year)), size = 3) +
  #2018
  geom_segment(x = 'BVR', xend = 'FCR', y = 0.019, yend = 0.033, alpha = 0.5,
               color = "#F8766D", linetype = 3) +
  #2019
  geom_segment(x = 'BVR', xend = 'FCR', y = 0.011, yend = 0.033, alpha = 0.5,
               color = "#B79F00", linetype = 3) +
  #2020
  geom_segment(x = 'BVR', xend = 'FCR', y = 0.005, yend = 0.021, alpha = 0.5,
               color = "#00BA38", linetype = 3) +
  #2021
  geom_segment(x = 'BVR', xend = 'FCR', y = 0.00712359, yend = 0.01952869, alpha = 0.5,
               color = "#00BFC4", linetype = 3) +
  #2022
  geom_segment(x = 'BVR', xend = 'FCR', y = 0.020, yend = 0.029, alpha = 0.5,
               color = "#619CFF", linetype = 3) +
  #2023
  geom_segment(x = 'BVR', xend = 'FCR', y = 0.014, yend = 0.068, alpha = 0.5,
               color = "#F564E3", linetype = 3) +
  annotate(geom = 'text', x= 1.5, y = 0.065, label = '*', size = 12) +
  guides(color = guide_legend(title = 'Year'), shape = guide_legend(title = 'Year'))

ComboFluxPlot <- ggarrange(CumulFePlot1, CumulMnPlot1, nrow = 1, common.legend = TRUE, legend = 'right')
```

```{r Figure 7, eval=TRUE, include=TRUE}
ComboFluxPlot
```

## Figure 8. Cumulative Sed Mass Plot
```{r Figure 8 script, eval=TRUE, include=FALSE}
plot_vline <- read_csv('~/Documents/R Projects/Sed Traps/TO_Dates.csv') %>% 
  mutate(Date = as.Date(Date),
         Event = as.factor(Event),
         Reservoir = as.factor(Reservoir),
         Layer = 'Hypolimnion')

FCRTO_vline <- plot_vline %>% 
  filter(Reservoir == 'FCR')

BVRTO_vline <- plot_vline %>% 
  filter(Reservoir == 'BVR')

strat_vline <- read_csv('StratStart2018_2023.csv') %>% 
  select(year, start, site_id, method) %>% 
  rename('Reservoir' = 'site_id',
         'Year' = 'year',
         'Date' = 'start') %>% 
  mutate(Layer = 'Hypolimnion')

HypCumulFluxPlot <- avgCumulFluxes2018_2023_long %>% 
  filter(Layer == 'Hypolimnion') %>%
  mutate(Metal = factor(Metal, levels = c('CumulFe_gm2', 'CumulMn_gm2'),
                        labels = c('Iron', 'Manganese'))) %>%
  ggplot() + theme_bw() + theme(panel.grid = element_blank()) + 
  facet_grid(cols = vars(Year), rows = vars(Metal), scales = 'free', space = "free_x") +
  scale_x_date(date_labels = '%b', breaks = '2 months', name = 'Date') +
  scale_y_continuous(name = 'Cumulative Mass Removal (g/m2)', limits = c(0, NA)) +
  geom_rect(aes(xmin = HOxOn, xmax = HOxOff, group = Group), ymin = -Inf, ymax = Inf, data = HOxBoxes, alpha = 0.2) + #creates the HOx boxes
  geom_vline(xintercept = FCRTO_vline$Date, linetype = 1, colour = '#D55E00') +
  geom_vline(xintercept = BVRTO_vline$Date, linetype = 2, colour = '#56B4E9') +
  geom_point(aes(as.Date(Date), Fluxes_gm2, colour = Reservoir, shape = Reservoir), size = 2)
```

```{r Figure 8, eval=TRUE, include=TRUE}
HypCumulFluxPlot
```

## Figure 9. 2023 and 2020 Comparison of 9m TFe, SFe, TMn, SMn in FCR

```{r Figure 9 set up, eval=TRUE, include=FALSE}
metals2023 <- metals %>% 
  filter(Year == 2023,
         MnDiff > 0)

Mnissues <- metals %>% 
  filter(Year == 2023,
         MnDiff < 0) %>% 
  rename('test1' = 'TMn_mgL',
         'test2' = 'SMn_mgL') %>% 
  rename('TMn_mgL' = 'test2',
         'SMn_mgL' = 'test1') %>% 
  mutate(MnDiff = TMn_mgL - SMn_mgL)

metals2023 <- rbind(metals2023, Mnissues) %>% 
  mutate(STFe = SFe_mgL/TFe_mgL,
         STMn = SMn_mgL/TMn_mgL) %>% 
  mutate(TO = NA,
         TO = ifelse(Year == 2023 & Reservoir == 'FCR' & Date < as.Date('2023-11-02'), 'Pre', TO),
         TO = ifelse(Year == 2023 & Reservoir == 'BVR' & Date < as.Date('2023-11-02'), 'Pre', TO),
         TO = ifelse(Year == 2023 & Reservoir == 'FCR' & Date >= as.Date('2023-11-02'), 'Post', TO),
         TO = ifelse(Year == 2023 & Reservoir == 'BVR' & Date >= as.Date('2023-11-02'), 'Post', TO))

metals2020 <- metals %>% 
  filter(Year == 2020,
         MnDiff > 0)

Mnissues20 <- metals %>% 
  filter(Year == 2020,
         MnDiff < 0) %>% 
  rename('test1' = 'TMn_mgL',
         'test2' = 'SMn_mgL') %>% 
  rename('TMn_mgL' = 'test2',
         'SMn_mgL' = 'test1') %>% 
  mutate(MnDiff = TMn_mgL - SMn_mgL)

metals2020 <- rbind(metals2020, Mnissues20) %>% 
  mutate(STFe = SFe_mgL/TFe_mgL,
         STMn = SMn_mgL/TMn_mgL) %>% 
  mutate(TO = NA,
         TO = ifelse(Year == 2020 & Reservoir == 'FCR' & Date < as.Date('2020-11-02'), 'Pre', TO),
         TO = ifelse(Year == 2020 & Reservoir == 'BVR' & Date < as.Date('2020-11-02'), 'Pre', TO),
         TO = ifelse(Year == 2020 & Reservoir == 'FCR' & Date >= as.Date('2020-11-02'), 'Post', TO),
         TO = ifelse(Year == 2020 & Reservoir == 'BVR' & Date >= as.Date('2020-11-02'), 'Post', TO),)
```

```{r Figure 9 script, eval=TRUE, include=FALSE}
Fe23 <- metals2023 %>% 
  select(Reservoir, Date, Site, Layer, Year, Depth_m, TFe_mgL, SFe_mgL, TO) %>%
  rename('Total' = 'TFe_mgL', 'Soluble' = 'SFe_mgL') %>% 
  pivot_longer(cols = 7:8, names_to = 'Type', values_to = 'Fe_mgL') %>% 
  filter(Layer == 'Hypolimnion')

Mn23 <- metals2023 %>% 
  select(Reservoir, Date, Site, Layer, Year, Depth_m, TMn_mgL, SMn_mgL, TO) %>%
  rename('Total' = 'TMn_mgL', 'Soluble' = 'SMn_mgL') %>% 
  pivot_longer(cols = 7:8, names_to = 'Type', values_to = 'Mn_mgL') %>% 
  filter(Layer == 'Hypolimnion') %>% 
  mutate(Type = factor(Type, levels = c('Soluble', 'Total')))

Fe20 <- metals2020 %>% 
  select(Reservoir, Date, Site, Layer, Year, Depth_m, TFe_mgL, SFe_mgL, TO) %>%
  rename('Total' = 'TFe_mgL', 'Soluble' = 'SFe_mgL') %>% 
  pivot_longer(cols = 7:8, names_to = 'Type', values_to = 'Fe_mgL') %>% 
  filter(Layer == 'Hypolimnion')

Mn20 <- metals2020 %>% 
  select(Reservoir, Date, Site, Layer, Year, Depth_m, TMn_mgL, SMn_mgL, TO) %>%
  rename('Total' = 'TMn_mgL', 'Soluble' = 'SMn_mgL') %>% 
  pivot_longer(cols = 7:8, names_to = 'Type', values_to = 'Mn_mgL') %>% 
  filter(Layer == 'Hypolimnion')

Mn23Reg <- Mn23 %>% 
  filter(Type == 'Total' & Reservoir == 'FCR' & Depth_m == 9 & TO == 'Pre' & Date >= as.Date('2023-03-23'))

Mn20Reg <- Mn20 %>% 
  filter(Type == 'Total' & Reservoir == 'FCR' & Depth_m == 9 & TO == 'Pre' & Date >= as.Date('2020-03-11'))

HOx2023 <- HOxBoxes %>% 
  filter(Year == 2023)

HOx2020 <- HOxBoxes %>% 
  filter(Year == 2020)

#Linear regression lines for release from sediment
Coef20 <- coef(lm(Mn20Reg$Mn_mgL ~ Mn20Reg$Date))[2]
Int20 <- coef(lm(Mn20Reg$Mn_mgL ~ Mn20Reg$Date))[1]

Coef23 <- coef(lm(Mn23Reg$Mn_mgL ~ Mn23Reg$Date))[2]
Int23 <- coef(lm(Mn23Reg$Mn_mgL ~ Mn23Reg$Date))[1]

Mod20 <- lm(Mn20Reg$Mn_mgL ~ Mn20Reg$Date)
summary(Mod20)
Mod23 <- lm(Mn23Reg$Mn_mgL ~ Mn23Reg$Date)
summary(Mod23)

#plots for Figure 9
STFe1 <- Fe20 %>% 
  filter(Layer == 'Hypolimnion' & Depth_m == 9 & Reservoir == 'FCR', Year == 2020) %>% 
ggplot() + 
  theme_bw(base_size = 18) + theme(panel.grid = element_blank(), legend.title = element_blank()) + 
  scale_x_date(date_labels = '%b', breaks = '2 months') + 
  scale_y_continuous(limits = c(0,10.5), name = 'Fe (mg/L)') + 
  geom_rect(aes(xmin = HOxOn, xmax = HOxOff, group = Group), ymin = -Inf, ymax = Inf, data = HOx2020, alpha = 0.2) +
  geom_vline(xintercept = as.Date('2020-11-02'), alpha = 0.75, linetype = 2) +
  geom_line(aes(as.Date(Date), Fe_mgL, color = as.factor(Type))) + 
  geom_point(aes(Date, Fe_mgL, shape = as.factor(Type), color = as.factor(Type)), size = 2) +
  ggtitle('2020') + #plot title
  theme(plot.title = element_text(hjust = 0.5)) + #centers title
  xlab(NULL)

STMn1 <- Mn20 %>% 
  filter(Layer == 'Hypolimnion' & Depth_m == 9 & Reservoir == 'FCR', Year == 2020) %>% 
  ggplot() + 
  theme_bw(base_size = 18) + theme(panel.grid = element_blank(), legend.title = element_blank()) + 
  scale_x_date(date_labels = '%b', breaks = '2 months') + 
  scale_y_continuous(limits = c(0,2.2), name = 'Mn (mg/L)') + 
  geom_rect(aes(xmin = HOxOn, xmax = HOxOff, group = Group), ymin = -Inf, ymax = Inf, data = HOx2020, alpha = 0.2) +
  geom_vline(xintercept = as.Date('2020-11-02'), alpha = 0.75, linetype = 2) +
  geom_line(aes(as.Date(Date), Mn_mgL, color = as.factor(Type))) +
  geom_smooth(aes(as.Date(Date), Mn_mgL), color = 'black', method = 'lm', formula = 'y~x', se = FALSE, data = Mn20Reg) +
  geom_point(aes(Date, Mn_mgL, shape = as.factor(Type), color = as.factor(Type)), size = 2) + xlab('Date') + geom_label(aes(x = as.Date('2020-04-20'), y = 2, label = 'Slope = 0.006 mg/Ld'))

STFe2 <- Fe23 %>% 
  filter(Layer == 'Hypolimnion' & Depth_m == 9 & Reservoir == 'FCR', Year == 2023) %>% 
  ggplot() + theme_bw(base_size = 18) + theme(panel.grid = element_blank(), legend.title = element_blank()) + scale_x_date(date_labels = '%b', breaks = '2 months') + 
  scale_y_continuous(limits = c(0,10.5), name = NULL) + 
  geom_rect(aes(xmin = HOxOn, xmax = HOxOff, group = Group), ymin = -Inf, ymax = Inf, data = HOx2023, alpha = 0.2) +
  geom_vline(xintercept = as.Date('2023-11-02'), alpha = 0.75, linetype = 2) +
  geom_line(aes(as.Date(Date), Fe_mgL, color = as.factor(Type))) + 
  geom_point(aes(Date, Fe_mgL, shape = as.factor(Type), color = as.factor(Type)), size = 2) +
  ggtitle('2023') + #plot title
  theme(plot.title = element_text(hjust = 0.5)) + #centers title
  xlab(NULL)

STMn2 <- Mn23 %>% 
  filter(Layer == 'Hypolimnion' & Depth_m == 9 & Reservoir == 'FCR', Year == 2023) %>% 
  ggplot() + theme_bw(base_size = 18) + theme(panel.grid = element_blank(), legend.title = element_blank()) + scale_x_date(date_labels = '%b', breaks = '2 months') + 
  scale_y_continuous(limits = c(0,2.2), name = NULL) + 
  geom_rect(aes(xmin = HOxOn, xmax = HOxOff, group = Group), ymin = -Inf, ymax = Inf, data = HOx2023, alpha = 0.2) +
  geom_vline(xintercept = as.Date('2023-11-02'), alpha = 0.75, linetype = 2) +
  geom_smooth(aes(as.Date(Date), Mn_mgL), color = 'black',
              method = 'lm', formula = 'y~x', se = FALSE, data = Mn23Reg) +
  geom_line(aes(as.Date(Date), Mn_mgL, color = as.factor(Type))) + 
  geom_point(aes(Date, Mn_mgL, shape = as.factor(Type), color = as.factor(Type)), size = 2) + xlab('Date') + geom_label(aes(x = as.Date('2023-04-20'), y = 2, label = 'Slope = 0.008 mg/Ld'))

comboPlot1 <- ggarrange(STFe1, STFe2, STMn1, STMn2, ncol = 2, nrow = 2, common.legend = TRUE, legend = 'bottom')
```

```{r Figure 9, eval=TRUE, include=TRUE}
comboPlot1
```

## Figure S1. Flux Plot

```{r Flux Plot Figure S1 script, eval=TRUE, include=FALSE}
avgFluxes2018_2023_long <-  avgFluxes2018_2023 %>% 
  pivot_longer(cols = c(7:8), values_to = 'Fluxes_gm2d', names_to = 'Metal')

FluxPlot <- avgFluxes2018_2023_long %>% 
  filter(Layer == 'Hypolimnion') %>% 
  mutate(Metal = factor(Metal, levels = c('AvgTFeFlux_gm2d', 'AvgTMnFlux_gm2d'),
                        labels = c('Iron', 'Manganese'))) %>% 
  ggplot() + theme_bw() + theme(panel.grid = element_blank()) + 
  facet_grid(cols = vars(Year), rows = vars(Metal), scales = 'free', space = "free_x") +
  scale_x_date(date_labels = '%b', breaks = '2 months') +
  scale_y_continuous(name = 'Sedimentation Fluxes (g/m2d)') +
  geom_rect(aes(xmin = HOxOn, xmax = HOxOff, group = Group), ymin = -Inf, ymax = Inf, data = HOxBoxes, alpha = 0.2) +
  geom_vline(xintercept = FCRTO_vline$Date, linetype = 1, colour = '#D55E00') +
  geom_vline(xintercept = BVRTO_vline$Date, linetype = 2, colour = '#56B4E9') +
  geom_line(aes(as.Date(Date), Fluxes_gm2d, colour = Reservoir)) +
  geom_point(aes(as.Date(Date), Fluxes_gm2d, colour = Reservoir, shape = Reservoir)) + xlab('Date')
```

```{r Figure S1, eval=TRUE, include=TRUE}
FluxPlot
```

## Figure S3. DO plot

```{r DO plot Figure S3 script, eval=TRUE, include=FALSE}
#DO Plot

DO1 <- FCRDO %>%
  mutate(DateTime = as.POSIXct(DateTime, format = "%Y-%m-%d %H:%M:%S")) %>% 
  select(Reservoir, Site, DateTime, RDO_mgL_9_adjusted) %>% 
  mutate(Date = date(DateTime)) %>% 
  group_by(Reservoir, Site, Date) %>% 
  summarise(DO_mgL_9m = mean(RDO_mgL_9_adjusted, na.rm = TRUE)) %>% 
  mutate(across(c('DO_mgL_9m'), ~ifelse(is.nan(.), NA, .)),
         Year = year(Date)) %>% 
  pivot_longer(cols = 4, names_to = 'Depth_m', values_to = 'DO_mgL') %>%
  mutate(Depth_m = factor(Depth_m, levels = c('DO_mgL_9m'),
                          labels = c('9')),
         Method = 'edi.271.9') %>% 
  drop_na()

DO2 <- BVRDO1 %>% 
  select(Reservoir, Site, DateTime, RDO_mgL_13) %>% 
  mutate(Date = date(DateTime)) %>% 
  group_by(Reservoir, Site, Date) %>% 
  summarise(DO_mgL_13m = mean(RDO_mgL_13, na.rm = TRUE)) %>% 
  mutate(across(c('DO_mgL_13m'), ~ifelse(is.nan(.), NA, .)),
         Year = year(Date)) %>% 
  pivot_longer(cols = 4, names_to = 'Depth_m', values_to = 'DO_mgL') %>%
  mutate(Depth_m = factor(Depth_m, levels = c('DO_mgL_13m'),
                          labels = c('13')),
         Method = 'edi.725.5') %>% 
  drop_na()

DO3 <- BVRDO2 %>% 
  filter(DateTime >= as.Date('2018-01-01')) %>% 
  select(Reservoir, Site, DateTime, MiniDotDO_mgL_10) %>% 
  mutate(Date = date(DateTime)) %>% 
  group_by(Reservoir, Site, Date) %>% 
  summarise(DO_mgL_10m = mean(MiniDotDO_mgL_10, na.rm = TRUE)) %>% 
  mutate(across(c('DO_mgL_10m'), ~ifelse(is.nan(.), NA, .)),
         Year = year(Date)) %>% 
  pivot_longer(cols = 4, names_to = 'Depth_m', values_to = 'DO_mgL') %>%
  mutate(Depth_m = factor(Depth_m, levels = c('DO_mgL_10m'),
                          labels = c('10')),
         Method = 'edi.725.5') %>% 
  drop_na()

DO4 <- Handhelds %>% 
  mutate(Date = date(DateTime),
         Year = year(DateTime)) %>% #HOx activated on 2018-04-23 @ noon
  filter(Year == 2018| Year == 2019 | Year == 2020,
         Reservoir == 'BVR' & Depth_m == 10 & 
           between(Date, as.Date('2018-05-04'), as.Date('2018-12-05'))|
           Reservoir == 'BVR' & Depth_m == 10 & 
           between(Date, as.Date('2019-05-18'), as.Date('2020-03-14'))|
           Reservoir == 'BVR' & Depth_m == 10 & 
           between(Date, as.Date('2020-05-16'), as.Date('2020-06-17'))|
           Reservoir == 'FCR' & Depth_m == 9 & Date < as.Date('2018-08-19')) %>% 
  select(Reservoir, Site, Date, Year, Depth_m, DO_mgL) %>% 
  mutate(Depth_m = factor(Depth_m),
         Method = 'edi.198.13') %>% 
  drop_na()

DO <- DO1 %>% 
  full_join(DO2) %>% 
  full_join(DO3) %>%
  full_join(DO4) %>% 
  filter(Date <= as.Date('2023-12-31')) %>%
  mutate(Method = factor(Method, levels = c('edi.271.9', 'edi.725.5', 'edi.198.13'))) %>%
  ungroup() %>% 
  #add_row(Reservoir = 'BVR', Site = 50, Date = as.Date('2018-11-17'),
  #        Year = 2018, Depth_m = as.factor(10), DO_mgL = NA, 
  #        Method = as.factor('edi.198.13')) %>%
  add_row(Reservoir = 'FCR', Site = 50, Date = as.Date('2018-08-19'),
          Year = 2018, Depth_m = as.factor(9), DO_mgL = NA, 
          Method = as.factor('edi.198.13')) %>%
  add_row(Reservoir = 'BVR', Site = 50, Date = as.Date('2018-12-05'),
          Year = 2018, Depth_m = as.factor(10), DO_mgL = NA, 
          Method = as.factor('edi.198.13')) %>%
  add_row(Reservoir = 'BVR', Site = 50, Date = as.Date('2019-05-17'),
          Year = 2019, Depth_m = as.factor(10), DO_mgL = NA, 
          Method = as.factor('edi.198.13')) %>%
  #add_row(Reservoir = 'BVR', Site = 50, Date = as.Date('2019-10-18'),
  #        Year = 2019, Depth_m = as.factor(10), DO_mgL = NA, 
  #        Method = as.factor('edi.198.13')) %>%
  add_row(Reservoir = 'BVR', Site = 50, Date = as.Date('2020-01-01'),
          Year = 2020, Depth_m = as.factor(10), DO_mgL = NA, 
          Method = as.factor('edi.725.5')) %>% 
  add_row(Reservoir = 'BVR', Site = 50, Date = as.Date('2020-03-15'),
          Year = 2020, Depth_m = as.factor(10), DO_mgL = NA, 
          Method = as.factor('edi.198.13')) %>% 
    add_row(Reservoir = 'BVR', Site = 50, Date = as.Date('2020-05-15'),
          Year = 2020, Depth_m = as.factor(10), DO_mgL = NA, 
          Method = as.factor('edi.198.13')) %>%
  add_row(Reservoir = 'BVR', Site = 50, Date = as.Date('2020-06-05'),
          Year = 2020, Depth_m = as.factor(10), DO_mgL = NA, 
          Method = as.factor('edi.198.13')) %>%
arrange(Date)

DO_plot1 <- DO %>% 
  ggplot() +
  scale_x_date(name = '', date_breaks = '1 year', date_labels = '%Y') +
  theme_bw(base_size = 18) + theme(panel.grid = element_blank()) +
  geom_line(aes(x = Date, y = DO_mgL, color = Reservoir, linetype = Method)) +
  scale_linetype_manual(values = c('solid', 'solid', 'dotted')) + 
  guides(linetype = guide_legend(title = 'Data Package')) +
  ylab('Dissolved Oxygen (mg/L)')
```

```{r Figure S3, eval=TRUE, include=TRUE}
DO_plot1
```

## Figure S4. DO on HOx Activation Day Plot
```{r DO plot figure S4 script, eval=TRUE, include=FALSE}

HOxOnCumulFluxes2018_2023 <- avgFluxes2018_2023 %>% 
  filter(TO == 'Pre' & Layer == 'Hypolimnion') %>%
  filter(Year == 2018 & between(Date, as.Date('2018-04-23'), as.Date('2018-08-12')) | #only include dates where HOx
           Year == 2019 & between(Date, as.Date('2019-06-03'), as.Date('2019-11-20'))  | #was active
           Year == 2020 & between(Date, as.Date('2020-06-29'), as.Date('2020-12-02')) |
           Year == 2021 & between(Date, as.Date('2021-06-11'), as.Date('2021-12-06')) |
           Year == 2022 & between(Date, as.Date('2022-05-19'), as.Date('2022-12-12')) |
           Year == 2023 & between(Date, as.Date('2023-03-06'), as.Date('2023-12-04'))) %>% 
  group_by(Reservoir, Layer, Year) %>% 
  mutate(CumulFe_gm2 = cumsum(AvgTFeFlux_gm2d*Duration_days),
         CumulMn_gm2 = cumsum(AvgTMnFlux_gm2d*Duration_days),
         CumulDuration_days = cumsum(Duration_days))

HOxOnCumulFluxSummary <- HOxOnCumulFluxes2018_2023 %>%
  group_by(Reservoir, Layer, Year) %>% 
  summarise(CumulFe_gm2 = max(CumulFe_gm2),
            CumulMn_gm2 = max(CumulMn_gm2),
            Duration_days = sum(Duration_days),
            Date = max(Date)) %>% 
  mutate(FeFlux = round(CumulFe_gm2/Duration_days, digits = 3),
         MnFlux = round(CumulMn_gm2/Duration_days, digits = 3))

DO_CumulMetalMass <- HOxOnCumulFluxSummary %>%
  ungroup() %>% 
  filter(Reservoir == 'FCR',
         Year >= 2018) %>% 
  select(Reservoir, Year, FeFlux, MnFlux) %>% 
  full_join(DOps2)

DO_CumulMetal_Feplot <- DO_CumulMetalMass %>% 
  ggplot() + theme_bw(base_size = 18) +
  geom_point(aes(x = AvgDOsat_percent_9, y = FeFlux, color = as.factor(Year),
                 shape = as.factor(Year)), size = 3) +
  ylab('Seasonal Flux (g/m2d)') + xlab('') +
  guides(color = guide_legend(title = 'Year'), shape = guide_legend(title = 'Year')) +
  ggtitle('Iron')

DO_CumulMetal_Mnplot <- DO_CumulMetalMass %>% 
  ggplot() + theme_bw(base_size = 18) +
  geom_point(aes(x = AvgDOsat_percent_9, y = MnFlux, color = as.factor(Year),
                 shape = as.factor(Year)), size = 3) +
  ylab('') + xlab('') +
  guides(color = guide_legend(title = 'Year'), shape = guide_legend(title = 'Year')) +
  ggtitle('Manganese')

DO_CumulMetal_plot1 <- ggarrange(DO_CumulMetal_Feplot, DO_CumulMetal_Mnplot,
                                 nrow = 1, common.legend = TRUE, legend = 'right', 
                                labels = c('A', 'B'))
DO_CumulMetal_plotFinal <- annotate_figure(DO_CumulMetal_plot1, bottom = text_grob('Dissolved Oxygen at HOx Activation (% saturation)', size = 18, hjust = 0.53, vjust = -1))
```

```{r Figure S4, eval=TRUE, include=TRUE}
DO_CumulMetal_plotFinal
```

## Tables
```{r Tables script, eval=TRUE, include=FALSE}
AvgFluxSummary2018_2023 <- avgFluxes2018_2023 %>% 
  filter(TO == 'Pre') %>% 
  group_by(Reservoir, Layer) %>% 
  summarise(MeanFeFlux = round(mean(AvgTFeFlux_gm2d, na.rm = T), digits = 3),
            MeanMnFlux = round(mean(AvgTMnFlux_gm2d, na.rm = T), digits = 3),
            MedianFeFlux = round(median(AvgTFeFlux_gm2d, na.rm = T), digits = 3),
            MedianMnFlux = round(median(AvgTMnFlux_gm2d, na.rm = T), digits = 3),
            MinFeFlux = round(min(AvgTFeFlux_gm2d, na.rm = T), digits = 3),
            MaxFeFlux = round(max(AvgTFeFlux_gm2d, na.rm = T), digits = 3),
            MinMnFlux = round(min(AvgTMnFlux_gm2d, na.rm = T), digits = 3),
            MaxMnFlux = round(max(AvgTMnFlux_gm2d, na.rm = T), digits = 3)) %>% 
  ungroup()

metals_HOx <- metals %>%
  filter(Layer == 'Hypolimnion') %>% 
  mutate(HOx = 'Off',
         HOx = if_else(Year == 2018 & Reservoir == 'FCR' &
                         between(Date, as.Date('2018-04-23'), as.Date('2018-07-30')), 'On', HOx),
         HOx = if_else(Year == 2018 & Reservoir == 'FCR' &
                         between(Date, as.Date('2018-08-09'), as.Date('2018-09-10')), 'On', HOx),
         HOx = if_else(Year == 2019 & Reservoir == 'FCR' &
                         between(Date, as.Date('2019-06-03'), as.Date('2019-06-17')), 'On', HOx),
         HOx = if_else(Year == 2019 & Reservoir == 'FCR' &
                         between(Date, as.Date('2019-07-08'), as.Date('2019-07-19')), 'On', HOx),
         HOx = if_else(Year == 2019 & Reservoir == 'FCR' &
                         between(Date, as.Date('2019-08-05'), as.Date('2019-08-19')), 'On', HOx),
         HOx = if_else(Year == 2019 & Reservoir == 'FCR' &
                         between(Date, as.Date('2019-09-02'), as.Date('2019-12-01')), 'On', HOx),
         HOx = if_else(Year == 2020 & Reservoir == 'FCR' &
                         between(Date, as.Date('2020-06-29'), as.Date('2020-07-12')), 'On', HOx),
         HOx = if_else(Year == 2020 & Reservoir == 'FCR' &
                         between(Date, as.Date('2020-07-13'), as.Date('2020-07-22')), 'On', HOx),
         HOx = if_else(Year == 2020 & Reservoir == 'FCR' &
                         between(Date, as.Date('2020-07-23'), as.Date('2020-09-11')), 'On', HOx),
         HOx = if_else(Year == 2020 & Reservoir == 'FCR' &
                         between(Date, as.Date('2020-09-25'), as.Date('2020-12-02')), 'On', HOx),
         HOx = if_else(Year == 2021 & Reservoir == 'FCR' &
                         between(Date, as.Date('2021-06-11'), as.Date('2021-06-26')), 'On', HOx),
         HOx = if_else(Year == 2021 & Reservoir == 'FCR' &
                         between(Date, as.Date('2021-06-30'), as.Date('2021-07-08')), 'On', HOx),
         HOx = if_else(Year == 2021 & Reservoir == 'FCR' &
                         between(Date, as.Date('2021-07-12'), as.Date('2021-07-14')), 'On', HOx),
         HOx = if_else(Year == 2021 & Reservoir == 'FCR' &
                         between(Date, as.Date('2021-07-26'), as.Date('2021-09-06')), 'On', HOx),
         HOx = if_else(Year == 2021 & Reservoir == 'FCR' &
                         between(Date, as.Date('2021-09-14'), as.Date('2021-12-06')), 'On', HOx),
         HOx = if_else(Year == 2022 & Reservoir == 'FCR' &
                         between(Date, as.Date('2022-03-22'), as.Date('2022-03-24')), 'On', HOx),
         HOx = if_else(Year == 2022 & Reservoir == 'FCR' &
                         between(Date, as.Date('2022-03-29'), as.Date('2022-04-03')), 'On', HOx),
         HOx = if_else(Year == 2022 & Reservoir == 'FCR' &
                         between(Date, as.Date('2022-03-22'), as.Date('2022-03-24')), 'On', HOx),
         HOx = if_else(Year == 2022 & Reservoir == 'FCR' &
                         between(Date, as.Date('2022-05-19'), as.Date('2022-06-14')), 'On', HOx),
         HOx = if_else(Year == 2022 & Reservoir == 'FCR' &
                         between(Date, as.Date('2022-06-21'), as.Date('2022-08-26')), 'On', HOx),
         HOx = if_else(Year == 2022 & Reservoir == 'FCR' &
                         between(Date, as.Date('2022-08-29'), as.Date('2022-10-01')), 'On', HOx),
         HOx = if_else(Year == 2022 & Reservoir == 'FCR' &
                         between(Date, as.Date('2022-10-03'), as.Date('2022-12-12')), 'On', HOx),
         HOx = if_else(Year == 2023 & Reservoir == 'FCR' &
                         between(Date, as.Date('2023-03-06'), as.Date('2023-12-04')), 'On', HOx)) %>% 
  group_by(Reservoir, Layer, HOx) %>% 
  summarise(AvgSTFe = round(mean(SFe_mgL/TFe_mgL, na.rm = TRUE), digits = 3),
            AvgSTMn = round(mean(SMn_mgL/TMn_mgL, na.rm = TRUE), digits = 3)) %>% 
  ungroup()
```

```{r Tables, eval=TRUE, include=TRUE}
gt(AvgFluxSummary2018_2023)
gt(metals_HOx)
```

```{r Save all figures to folder if desired, eval=FALSE}

#Figure 6
ggsave(DO_MaxMetal_plotFinal, file = 'DOps_MaxMetal.png', path = myFilePath,
       height = 5, width = 8, units = 'in')
#Figure 7
ggsave(plot = ComboFluxPlot, filename = 'CumulHypFlux_tTest_2018-2023.png', path = myFilePath, height = 4, width = 8, units = 'in')
#Figure 8
ggsave(plot = HypCumulFluxPlot, filename = 'CumulHypFlux_HOxBoxes_2018-2023.png', path = myFilePath, height = 4, width = 8, units = 'in')
#Figure 9
ggsave(comboPlot1, file = 'STFeMn_2020_2023.png', path = myFilePath, height = 6, width = 8, units = 'in')
#Figure S1
ggsave(plot = FluxPlot, filename = 'Flux_HOxBoxes_2018-2023.png', path = myFilePath, height = 4, width = 8, units = 'in')
#Figure S3
ggsave(DO_plot1, file = 'HypolimnionDO_2018_2023.png', path = myFilePath, height = 4, 
       width = 6, units = 'in')
#Figure S4
ggsave(DO_CumulMetal_plotFinal, file = 'DOps_MaxMetal.png', path = myFilePath,
       height = 5, width = 8, units = 'in')
```