### 2025-08-28 Weather_com_figure_ABP.Rmd
Updated the script so each plot is seperate

---
title: "Figure for Weather Comparison"
output: html_notebook
author: "Adrienne Breef-Pilz"
date: "2025-08-27"
---

This Markdown makes a figure comparing the average temperature, daily rain fall, and windspeed from Falling Creek Reservoir from 2018-2023. It uses data from EDI. The plot is for the SI of Cece Wood's Sed Trap manuscript.

```{r set up}
# load data packages
pacman::p_load(tidyverse)

```

```{r read in data}

# read in data from EDI and filter down to 2018-2023 and variables of temp, rainfall, and wind speed

# reading in the data package from 2024 to stay consistent with the other data products
met <- read_csv("https://pasta.lternet.edu/package/data/eml/edi/389/8/d4c74bbb3b86ea293e5c52136347fbb0")

# filter down the data frame to the correct date and select the columns we need
met_2018_23 <- met|>
  filter(DateTime>=as.Date("2018-01-01") & DateTime< as.Date("2024-01-01"))|>
  select(DateTime, AirTemp_C_Average, Rain_Total_mm, WindSpeed_Average_m_s)|>
  mutate(Date = as.Date(DateTime))

```

```{r Daily average temp and wind speed}
# Get the daily average of temperature and wind speed

met_avg <- met_2018_23|>
  select(Date, AirTemp_C_Average, WindSpeed_Average_m_s)|>
  group_by(Date)|>
  summarise_at(vars(AirTemp_C_Average:WindSpeed_Average_m_s), mean, na.rm = TRUE)

```

```{r Daily sum of rain}
#Get the daily sum of rain fall and convert it from mm to cm

met_sum <- met_2018_23|>
  select(Date, Rain_Total_mm)|>
  group_by(Date)|>
  summarise(Rain_Total_sum_cm = sum(Rain_Total_mm)/10)

```

```{r merge}
# Merge the daily average data frame with the daily sum of rainfall data frame
met_all <- merge(met_avg, met_sum)

met_all <- met_all|>
  mutate(DOY = yday(Date),
         Year = year(Date),
         ddate = as.Date(DOY-1)) # make the DOY into a date but it is off by 1 so subtract and then convert it into a date format. This way they all have the same year and will stack on top of one another

# pivot the data frame longer to use facet_grid below
met_all_long <- met_all|>
  pivot_longer(!c(Date, DOY, Year, ddate), names_to = "variable", values_to = "obs")

```

```{r Figure}
# Make the figures on one panel and save it in the figures folder
# Just going to comment this out, but save it

# # Rename the variable to fit in the labels
# hum_names <- as_labeller(
#      c(`Rain_Total_sum_cm` = "Daily Rain Toal cm", `AirTemp_C_Average` = "Daily Average Air Temp C",`WindSpeed_Average_m_s` = "Daily Averge Wind Speed m/s"))
# 
# # Make the plot here
# ggplot(met_all_long, aes(x = ddate, y = obs, color = as.factor(Year)))+
#   geom_point()+ # point or line?
#   facet_grid(vars(variable), scales = "free", labeller = hum_names)+
#   labs(color='Year') + 
#   ylab("")+
#   xlab("Date")+
#   scale_x_date(date_breaks = "2 month", date_minor_breaks = "1 month",
#              date_labels = "%B")+
#   theme_bw(base_size = 15) #size of all the text in the figure. 
# 
# 
#  # can change dimensions as you see fit 
# ggsave("figures/Met_com_2018_2023.png", width = 10, height = 8, dpi = 300, units = "in", device='png')
```
```{r Function to Make Plots}

# Function to make the Plots

met_plot <- function(data= met_all_long,
                     VAR, 
                     y_lab){

a <- met_all_long%>%
  filter(variable == VAR)%>%
ggplot(., aes(x = ddate, y = obs, color = as.factor(Year)))+
  geom_point()+ 
  labs(color='Year') +
  ylab(y_lab)+
  xlab("Date")+
  scale_x_date(date_breaks = "2 month", date_minor_breaks = "1 month",
             date_labels = "%B")+
  theme_bw(base_size = 18) #size of all the text in the figure.

return(a)
}
```

```{r Temp Plot}
temp <- met_plot(VAR = "AirTemp_C_Average", y_lab = expression(""* ~degree*C))
temp

 # can change dimensions as you see fit
ggsave("figures/SI_Met_Comp/Temp_2018_2023.png", plot = temp,width = 10, height = 8, dpi = 300, units = "in", device='png')

```

```{r Wind speed}

wind <- met_plot(VAR = "WindSpeed_Average_m_s", y_lab = "m/s")
wind

 # can change dimensions as you see fit
ggsave("figures/SI_Met_Comp/Wind_2018_2023.png", plot = wind, width = 10, height = 8, dpi = 300, units = "in", device='png')

```
```{r Rain}
rain <- met_plot(VAR = "Rain_Total_sum_cm", y_lab = "centimeters")
rain

 # can change dimensions as you see fit
ggsave("figures/SI_Met_Comp/Rain_2018_2023.png", plot = rain, width = 10, height = 8, dpi = 300, units = "in", device='png')

```

```{r Violin plots}
# Playing with box plots

Atemp <- met_2018_23%>%
  mutate(Year = year(DateTime))%>%
  ggplot(., aes(x = Year, y = AirTemp_C_Average, group = Year, fill = as.factor(Year)))+
  geom_violin()+
  theme_bw()

Atemp

Wspeed <- met_2018_23%>%
  mutate(Year = year(DateTime))%>%
  ggplot(., aes(x = Year, y = WindSpeed_Average_m_s, group = Year, fill = as.factor(Year)))+
  geom_violin()+
  theme_bw()

Wspeed

```